---
title: "Visualizing Circular and Directional Data with ggcircular"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualizing Circular and Directional Data with ggcircular}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 5)
```

# Abstract

This demonstration presents a reproducible workflow for simulated animal
movement data. The aim is to show how rose diagrams, circular density estimates
and mean directions can be combined in a `ggplot2` grammar.

# Introduction

Directional variables occur in wind, movement, orientation and time-of-day
data. Their periodic nature requires summaries and graphics that respect the
circle.

# Circular data and specialized graphics

Circular graphics should keep the endpoints of the angular scale adjacent.
They should also distinguish directionality, concentration and uncertainty.

# The grammar of ggcircular

```{r}
library(ggplot2)
library(dplyr)
library(ggcircular)
```

# Simulated trajectories

```{r}
ggplot(animal_steps, aes(x = x, y = y, group = id, colour = id)) +
  geom_path(alpha = 0.7) +
  coord_equal() +
  theme_minimal()
```

# Directional features

```{r}
movement_summary <- animal_steps |>
  group_by(state) |>
  summarise(
    mean_step = mean(step_length, na.rm = TRUE),
    median_step = median(step_length, na.rm = TRUE),
    .groups = "drop"
  )

movement_summary
```

# Bearings by state

```{r}
ggplot(animal_steps, aes(x = bearing, fill = state)) +
  geom_rose(bins = 24, alpha = 0.7) +
  facet_wrap(~ state) +
  scale_x_circular_degrees() +
  coord_circular() +
  theme_circular()
```

# Turn angles by state

```{r}
ggplot(animal_steps, aes(x = turn_angle, fill = state)) +
  geom_rose(bins = 24, alpha = 0.7) +
  geom_mean_direction() +
  facet_wrap(~ state) +
  scale_x_circular_radians() +
  coord_circular() +
  theme_circular()
```

# Mean directions and resultant length

```{r}
animal_steps |>
  group_by(state) |>
  circular_summary(turn_angle)
```

# Circular densities

```{r}
ggplot(animal_steps, aes(x = turn_angle, colour = state)) +
  geom_circular_density(linewidth = 1) +
  scale_x_circular_radians() +
  coord_circular() +
  theme_circular()
```

# Interpretation

In the simulated data, directed movement has longer steps and more concentrated
turn angles. Encamped movement has shorter steps and more diffuse turning.
Exploratory movement is intermediate. These statements describe the simulated
dataset and should not be generalized beyond it.

# Discussion

The example shows how the same grammar can summarize raw angular observations,
derived movement quantities and grouped comparisons.

# Reproducibility

All data used here are included in the package as simulated datasets. The code
chunks can be evaluated after installing `ggcircular`.
