---
title: "Axial data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Axial data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 5)
```

# Directional versus axial

Directional observations have a head and tail. Axial observations represent an
orientation where angles separated by `pi` are equivalent.

# Examples

Fiber orientations, fault orientations and undirected line segments are common
examples of axial data.

# Doubling angles

The usual computational approach doubles angles, applies directional circular
statistics, then halves the resulting mean direction.

```{r}
library(ggplot2)
library(ggcircular)

circular_summary(axial_orientations, orientation, axial = TRUE)
```

# Axial rose diagram

```{r}
ggplot(axial_orientations, aes(x = orientation, fill = group)) +
  geom_rose(bins = 18, axial = TRUE, alpha = 0.7) +
  scale_x_circular_degrees(limits = c(0, pi)) +
  coord_circular() +
  theme_circular()
```

# Axial mean

```{r}
ggplot(axial_orientations, aes(x = orientation, fill = group)) +
  geom_rose(bins = 18, axial = TRUE, alpha = 0.5) +
  geom_mean_direction(axial = TRUE) +
  scale_x_circular_degrees(limits = c(0, pi)) +
  coord_circular() +
  theme_circular()
```

# Axial density

```{r}
ggplot(axial_orientations, aes(x = orientation, colour = group)) +
  geom_circular_density(axial = TRUE, linewidth = 1) +
  scale_x_circular_degrees(limits = c(0, pi)) +
  coord_circular() +
  theme_circular()
```

# Communication

Reports should state clearly whether angles are directional or axial, because
the interpretation of the same numeric angle can change.
