---
title: "Circular distributions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Circular distributions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 5)
```

# Uniform circular distribution

The circular uniform density is constant over the angular period.

```{r}
library(ggplot2)
library(ggcircular)

ggplot() +
  stat_uniform_circular() +
  scale_x_circular_degrees() +
  coord_circular() +
  theme_circular()
```

# Von Mises distribution

The von Mises distribution is a common circular analogue of a unimodal normal
model. The parameter `mu` controls the mean direction and `kappa` controls
concentration.

```{r}
ggplot() +
  stat_vonmises(mu = pi / 2, kappa = 3, linewidth = 1) +
  scale_x_circular_degrees() +
  coord_circular() +
  theme_circular()
```

# Overlay a theoretical density

```{r}
ggplot(wind_directions, aes(x = direction)) +
  geom_rose(aes(y = after_stat(density)), bins = 24, alpha = 0.5) +
  stat_vonmises(mu = pi / 2, kappa = 3, linewidth = 1) +
  scale_x_circular_degrees() +
  coord_circular() +
  theme_circular()
```

# Simple fitted von Mises density

```{r}
ggplot(wind_directions, aes(x = direction)) +
  geom_rose(aes(y = after_stat(density)), bins = 24, alpha = 0.5) +
  stat_vonmises_fit(linewidth = 1) +
  scale_x_circular_degrees() +
  coord_circular() +
  theme_circular()
```

# Empirical and theoretical comparison

The fitted density is a descriptive approximation. It is useful as a visual
reference and should not replace formal model checking when a parametric model
is used for inference.
