---
title: "Circular-circular workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Circular-circular workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4
)
```

This vignette shows a basic circular-circular workflow. Both variables are
angles, so the sample space is a torus.

```{r}
library(ggplot2)
library(ggcircular)

dat <- simulate_tor_diagnostic(
  n = 260,
  scenario = "diagonal",
  seed = 2
)
```

The toroidal flow displays conditional binned mass between angular sectors.
The toroidal topography estimates `f(phi | theta)`. The conditional ridge
extracts a dominant modal relation.

```{r}
plot_toroidal_flow(dat$theta, dat$phi, n_sectors = 18)
```

```{r}
plot_toroidal_topography(dat$theta, dat$phi, n_theta = 50, n_phi = 50, conditional = TRUE)
```

```{r}
plot_toroidal_ridge(dat$theta, dat$phi, n_theta = 50, n_phi = 50)
```

The same displays can be built as `ggplot2` layers.

```{r}
ggplot(dat, aes(x = theta, y = phi)) +
  stat_toroidal_topography(n_theta = 50, n_phi = 50, conditional = TRUE) +
  stat_toroidal_ridge(n_theta = 50, n_phi = 50, linewidth = 1) +
  coord_equal()
```

Ridges are most interpretable when a dominant mode is meaningful. For
bifurcating conditional densities, use the topography as the primary display and
interpret a single ridge cautiously.
