---
title: "Getting started with ggcircular"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with ggcircular}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 5)
```

# Overview

`ggcircular` extends `ggplot2` for circular, axial and directional data. This
vignette gives a complete first tour of the package: angle conventions, rose
diagrams, circular densities, mean directions, uncertainty, axial data, movement
data, mixtures of von Mises distributions and model diagnostics.

# Installation

Install the released version from CRAN with `install.packages("ggcircular")`.
Install the development version from GitHub when you need unreleased features.

```{r}
library(ggplot2)
library(dplyr)
library(ggcircular)
```

# Why circular data are different

Circular observations live on a periodic scale. In radians, `0` and `2 * pi`
represent the same direction. This means that linear tools can fail near the
boundary.

```{r}
boundary_angles <- tibble(
  theta = c(0.05, 0.10, 2 * pi - 0.10, 2 * pi - 0.05)
)

boundary_angles |>
  summarise(
    arithmetic_mean = mean(theta),
    circular_mean = mean_direction(theta),
    Rbar = mean_resultant_length(theta)
  )
```

The arithmetic mean is near `pi`, even though the observations are concentrated
near zero. The circular mean uses sine and cosine components, so it respects the
periodic scale.

# Data included in the package

The package ships with four simulated datasets. They are small enough for
examples and large enough to show realistic grouped workflows.

```{r}
glimpse(wind_directions)
glimpse(animal_steps)
glimpse(hourly_activity)
glimpse(axial_orientations)
```

```{r}
wind_directions |>
  count(season, station)
```

# Directional versus axial data

Directional data have an arrow. For example, a bearing of north and a bearing of
south are different directions. Axial data have an orientation but no arrow, so
an angle and the angle plus `pi` are equivalent.

```{r}
directional <- c(0, pi)
axial <- c(0, pi)

tibble(
  case = c("directional", "axial"),
  Rbar = c(
    mean_resultant_length(directional),
    mean_resultant_length(axial, axial = TRUE)
  ),
  mean = c(
    mean_direction(directional),
    mean_direction(axial, axial = TRUE)
  )
)
```

For axial calculations, `ggcircular` doubles the angles internally, computes the
directional statistic, then transforms the answer back to the original scale.

# Conventions for directions and bearings

The internal default unit is radians. Helpers are provided for degrees, hours
and compass labels.

```{r}
tibble(
  degrees = c(0, 90, 180, 270),
  radians = deg_to_rad(degrees),
  hours = rad_to_hour(radians),
  compass = rad_to_compass(radians)
)
```

Compass labels use the bearing convention: zero points north and angles increase
clockwise. Use this with `coord_circular(zero = "north", direction =
"clockwise")`.

For mathematical plots, the default coordinate convention is zero at east and
positive angles rotating counterclockwise. For axial data, set `axial = TRUE`
because `theta` and `theta + pi` represent the same orientation.

# First rose diagram

A rose diagram is a circular histogram. The first and last bins are adjacent on
the circle.

```{r}
ggplot(wind_directions, aes(x = direction)) +
  geom_rose(bins = 16) +
  scale_x_circular_degrees() +
  coord_circular() +
  theme_circular()
```

# Counts, densities and proportions

`geom_rose()` exposes computed variables such as `count`, `density` and
`proportion`. These are available with `after_stat()`.

```{r}
ggplot(wind_directions, aes(x = direction)) +
  geom_rose(
    aes(fill = after_stat(proportion)),
    bins = 16,
    normalize = "proportion"
  ) +
  scale_x_circular_degrees() +
  coord_circular() +
  theme_rose()
```

# Groups and facets

Because the layers follow the `ggplot2` grammar, standard grouping, colouring
and faceting workflows work naturally.

```{r}
ggplot(wind_directions, aes(x = direction, fill = season)) +
  geom_rose(bins = 16, alpha = 0.75) +
  facet_wrap(~ season) +
  scale_x_circular_degrees() +
  coord_circular() +
  theme_circular()
```

# Circular density

`geom_circular_density()` estimates a smooth density on the circle using a von
Mises kernel. The estimate wraps around the origin.

```{r}
ggplot(wind_directions, aes(x = direction)) +
  geom_rose(aes(y = after_stat(density)), bins = 24, alpha = 0.35) +
  geom_circular_density(linewidth = 1) +
  scale_x_circular_degrees() +
  coord_circular() +
  theme_circular()
```

The bandwidth can be adjusted. Smaller values show more local variation.

```{r}
ggplot(wind_directions, aes(x = direction)) +
  geom_circular_density(bw = 0.25, linewidth = 1) +
  geom_circular_density(bw = 0.75, linetype = 2) +
  scale_x_circular_degrees() +
  coord_circular() +
  theme_circular()
```

# Mean direction and resultant length

The mean resultant length `Rbar` measures concentration. Values close to one
indicate strong concentration; values close to zero indicate weak or cancelling
directionality.

```{r}
wind_directions |>
  group_by(season) |>
  circular_summary(direction)
```

```{r}
ggplot(wind_directions, aes(x = direction, colour = season)) +
  geom_circular_density(linewidth = 1) +
  geom_mean_direction(length = "resultant") +
  scale_x_circular_degrees() +
  coord_circular() +
  theme_circular()
```

# Uncertainty and circular tests

`circular_mean_ci()` computes large-sample or bootstrap intervals for the mean
direction. `rayleigh_test()` provides a basic test against circular uniformity.

```{r}
circular_mean_ci(wind_directions$direction, method = "large_sample")
rayleigh_test(wind_directions$direction)
```

```{r}
ggplot(wind_directions, aes(x = direction)) +
  geom_rose(bins = 16, alpha = 0.8) +
  stat_circular_test(test = "rayleigh", y = 1.1, size = 3) +
  scale_x_circular_degrees() +
  coord_circular() +
  theme_circular()
```

# Compass display

For bearing-like data, compass labels and geographic orientation are usually
easier to read.

```{r}
ggplot(wind_directions, aes(x = direction)) +
  geom_rose(bins = 16, aes(fill = after_stat(count))) +
  geom_mean_direction() +
  scale_x_circular_compass() +
  coord_circular(zero = "north", direction = "clockwise") +
  theme_compass()
```

# Axial data

Set `axial = TRUE` when orientations are modulo `pi`.

```{r}
axial_orientations |>
  group_by(group) |>
  circular_summary(orientation, axial = TRUE)
```

```{r}
ggplot(axial_orientations, aes(x = orientation, fill = group)) +
  geom_rose(bins = 18, axial = TRUE, alpha = 0.7) +
  geom_mean_direction(axial = TRUE) +
  scale_x_circular_degrees(limits = c(0, pi)) +
  coord_circular() +
  theme_circular()
```

# Movement data

Movement tracks naturally produce step lengths, bearings and turn angles.

```{r}
animal_steps |>
  group_by(state) |>
  summarise(
    mean_step = mean(step_length, na.rm = TRUE),
    median_step = median(step_length, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r}
ggplot(animal_steps, aes(x = x, y = y, group = id, colour = id)) +
  geom_path(alpha = 0.7) +
  coord_equal() +
  theme_minimal()
```

```{r}
plot_state_angles(animal_steps, angle = turn_angle, state = state, type = "rose")
```

```{r}
plot_state_angles(animal_steps, angle = turn_angle, state = state, type = "density")
```

# Mixtures of von Mises distributions

Mixtures provide a descriptive way to represent multimodal circular
distributions. The EM fit can depend on initialization, so use `seed`, `nstart`
and `glance_circular()` when reproducibility or convergence matters.

```{r}
fit_mix <- fit_vonmises_mixture(
  wind_directions$direction,
  k = 2,
  nstart = 3,
  seed = 2026,
  max_iter = 200
)

tidy_circular(fit_mix)
glance_circular(fit_mix)
```

```{r}
ggplot(wind_directions, aes(x = direction)) +
  geom_rose(aes(y = after_stat(density)), bins = 24, alpha = 0.35) +
  stat_vonmises_mixture(fit = fit_mix, linewidth = 1) +
  scale_x_circular_degrees() +
  coord_circular() +
  theme_circular()
```

# Angular model diagnostics

`ggcircular` provides basic methods for angular model summaries and residual
diagnostics. The example below uses a small mock object with the same observed
and fitted angle fields expected from supported angular model classes.

```{r}
fit <- structure(
  list(
    y = wind_directions$direction[1:50],
    mui = normalize_angle(wind_directions$direction[1:50] + rnorm(50, 0, 0.15)),
    term_labels = c("intercept", "speed")
  ),
  class = "angular"
)

tidy_circular(fit)
glance_circular(fit)
circular_model_diagnostics(fit)
```

```{r}
autoplot(fit, type = "residuals_density")
```

# Circular posterior draws

When the optional `posterior` package is installed, circular posterior draws can
be converted to a long format and summarized with circular statistics.

```{r}
if (requireNamespace("posterior", quietly = TRUE)) {
  set.seed(1)
  draws <- posterior::draws_df(
    theta = rnorm(400, mean = pi / 3, sd = 0.25),
    phi = rnorm(400, mean = pi, sd = 0.35)
  )

  circular_draws <- as_circular_draws(draws, variables = c("theta", "phi"))
  summarise_circular_draws(circular_draws)
}
```

```{r}
if (requireNamespace("posterior", quietly = TRUE)) {
  autoplot_circular_draws(circular_draws)
}
```

# Experimental features

The optional model integrations are experimental. They are intended to make
diagnostics easier for workflows built with `CircularRegression`, `momentuHMM`
and `posterior`, while keeping those packages in `Suggests`. The functions give
explicit errors when an optional package is required but not installed.

# Statistical limitations

Circular graphics are descriptive and should be read with the data-generating
context in mind.

1. A rose diagram can change visually with the number of bins.
2. A circular mean is unstable when `Rbar` is close to zero.
3. Directional and axial data require different summaries.
4. Compass bearings and mathematical angles use different zero directions.
5. Multimodal data should not be summarized only by one mean direction.
6. The automatic circular density bandwidth is a heuristic.
7. `estimate_kappa()` is a descriptive approximation, not a full uncertainty
   analysis.
8. Rayleigh and Watson-Williams tests have classical assumptions that should be
   checked before confirmatory use.

# Next steps

The focused vignettes go deeper into rose diagrams, circular density, mean
direction, axial data, movement data, theoretical distributions, angular model
diagnostics, and spherical or posterior workflows.
