---
title: "Getting started with ggcircular"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with ggcircular}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 5)
```

# Why circular data are different

Circular observations live on a periodic scale. In radians, `0` and `2 * pi`
represent the same direction. A linear histogram or arithmetic mean can be
misleading when observations concentrate near the boundary.

# Directional versus axial data

Directional data have an arrow: north and south are different directions.
Axial data have an orientation but no arrow: an angle and the angle plus `pi`
are equivalent. `ggcircular` exposes an `axial = TRUE` argument in the core
statistics and geoms. Internally, the usual calculation doubles the angles,
computes directional summaries, then divides the resulting angle by two.

# Why a standard histogram is not enough

The first and last bins of a circular histogram are adjacent on the circle. A
rose diagram respects this topology when combined with a polar coordinate
system.

# The ggplot2 logic

`ggcircular` layers are ordinary `ggplot2` layers. They can be added to plots,
faceted, grouped and combined with familiar scales and themes.

```{r}
library(ggplot2)
library(ggcircular)
```

# First rose diagram

```{r}
ggplot(wind_directions, aes(x = direction)) +
  geom_rose(bins = 16) +
  scale_x_circular_degrees() +
  coord_circular() +
  theme_circular()
```

# Add density and mean

```{r}
ggplot(wind_directions, aes(x = direction)) +
  geom_rose(aes(y = after_stat(density)), bins = 24, alpha = 0.4) +
  geom_circular_density(linewidth = 1) +
  geom_mean_direction() +
  scale_x_circular_degrees() +
  coord_circular() +
  theme_circular()
```

# Choose units

The package includes x scales for radians, degrees, hours and compass labels.
Compass labels are intended for use with `coord_circular(zero = "north",
direction = "clockwise")`.

```{r}
ggplot(wind_directions, aes(x = direction)) +
  geom_rose(bins = 16) +
  scale_x_circular_compass() +
  coord_circular(zero = "north", direction = "clockwise") +
  theme_compass()
```

# Go further

The remaining vignettes focus on rose diagrams, circular density estimation,
mean directions, axial data, movement data and theoretical circular
distributions.
