---
title: "Mean direction and uncertainty"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mean direction and uncertainty}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 5)
```

# Circular mean

The circular mean is computed from the mean cosine and sine components. This
avoids the boundary problem that affects arithmetic means.

```{r}
library(ggplot2)
library(dplyr)
library(ggcircular)

circular_summary(wind_directions, direction)
```

# Why arithmetic means fail

Angles close to `0` and `2 * pi` are close on the circle but far on the real
line. Component-based averages respect the circle.

# Resultant length

The mean resultant length `Rbar` is between zero and one. Values near one
indicate concentration near a common direction; values near zero indicate weak
or cancelling directionality.

# Mean by group

```{r}
wind_directions |>
  group_by(season) |>
  circular_summary(direction)
```

# Displaying the mean

```{r}
ggplot(wind_directions, aes(x = direction)) +
  geom_rose(bins = 24) +
  geom_mean_direction(length = "resultant") +
  scale_x_circular_degrees() +
  coord_circular() +
  theme_circular()
```

# Arcs of uncertainty

`stat_mean_direction()` computes approximate confidence limits when
`conf.int = TRUE`. `geom_confidence_arc()` can display any angular interval
stored as lower and upper endpoints.

```{r}
summaries <- wind_directions |>
  group_by(season) |>
  circular_summary(direction)

ggplot(summaries, aes(x = mean, y = Rbar)) +
  geom_point() +
  scale_x_circular_degrees() +
  coord_circular() +
  theme_circular()
```

# Uniform or nearly uniform cases

When `Rbar` is close to zero, the mean direction is unstable and should be
interpreted cautiously.
