---
title: "Movement data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Movement data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 5)
```

# Tracks and directions

Trajectories produce step lengths, bearings and turn angles. These quantities
are natural inputs for circular graphics.

```{r}
library(ggplot2)
library(dplyr)
library(ggcircular)
```

# Coordinates, steps, bearings and turns

```{r}
raw_tracks <- animal_steps |>
  select(id, time, x, y)

steps <- raw_tracks |>
  mutate_directional_features(x = x, y = y, id = id, time = time)
```

# Visualize a trajectory

```{r}
ggplot(animal_steps, aes(x = x, y = y, group = id)) +
  geom_path(alpha = 0.5) +
  geom_direction_arrow(aes(angle = bearing, length = step_length), alpha = 0.35) +
  coord_equal() +
  theme_minimal()
```

# Bearings by state

```{r}
ggplot(animal_steps, aes(x = bearing, fill = state)) +
  geom_rose(bins = 24) +
  facet_wrap(~ state) +
  scale_x_circular_degrees() +
  coord_circular() +
  theme_circular()
```

# Turn angles by state

```{r}
ggplot(animal_steps, aes(x = turn_angle, fill = state)) +
  geom_rose(bins = 24) +
  geom_mean_direction() +
  facet_wrap(~ state) +
  scale_x_circular_radians() +
  coord_circular() +
  theme_circular()
```

# Comparison by state

```{r}
ggplot(animal_steps, aes(x = turn_angle, colour = state)) +
  geom_circular_density(linewidth = 1) +
  scale_x_circular_radians() +
  coord_circular() +
  theme_circular()
```

# Links with HMM and SSF workflows

The same summaries can be applied to observed states, hidden states or
step-selection covariates. The package intentionally keeps model dependencies
out of the core package so these workflows remain flexible.
