---
title: "Smoothing and support diagnostics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Smoothing and support diagnostics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4
)
```

Conditional displays depend on smoothing parameters. This vignette shows how to
choose them by cross-validated conditional likelihood and how to flag angular
regions with weak marginal support.

```{r}
library(ggcircular)

cyl <- simulate_cyl_diagnostic(n = 180, scenario = "heteroscedastic", seed = 3)
tor <- simulate_tor_diagnostic(n = 200, scenario = "doubling", seed = 4)
```

For circular-linear data, select both the circular concentration and the linear
bandwidth.

```{r}
cyl_smoothing <- select_cyl_smoothing(
  theta = cyl$theta,
  x = cyl$x,
  kappa_values = c(8, 14),
  h_values = c(0.25, 0.45),
  n_folds = 3,
  seed = 5
)

head(cyl_smoothing)
```

For circular-circular data, select the conditioning and response concentrations.

```{r}
tor_smoothing <- select_toroidal_smoothing(
  theta = tor$theta,
  phi = tor$phi,
  kappa_theta_values = c(8, 14),
  kappa_phi_values = c(8, 14),
  n_folds = 3,
  seed = 6
)

head(tor_smoothing)
```

The support diagnostic should be inspected before overinterpreting conditional
features in sparsely sampled angular regions.

If `R(theta)` is the kernel support relative to the best-supported direction,
the local standard error is approximately inflated by `1 / sqrt(R(theta))`
when other ingredients are fixed. Thus a threshold of 0.25 flags regions with
about twice the best-supported standard error, while the default 0.15 permits
an inflation of about 2.58. Use 0.15 for an exploratory display and 0.25 for a
more conservative reading. If the substantive interpretation changes between
these thresholds, report the sensitivity rather than selecting one value
silently.

```{r}
support <- marginal_support_data(
  theta = cyl$theta,
  kappa = cyl_smoothing$kappa[1],
  relative_threshold = 0.20
)

head(support)
plot_marginal_support(
  theta = cyl$theta,
  kappa = cyl_smoothing$kappa[1],
  relative_threshold = 0.20
)
```

```{r}
support_default <- marginal_support_data(
  theta = cyl$theta,
  kappa = cyl_smoothing$kappa[1],
  relative_threshold = 0.15
)
support_conservative <- marginal_support_data(
  theta = cyl$theta,
  kappa = cyl_smoothing$kappa[1],
  relative_threshold = 0.25
)

c(
  default_fraction_flagged = mean(support_default$low_support),
  conservative_fraction_flagged = mean(support_conservative$low_support)
)
```
