---
title: "Spherical directions and circular posterior draws"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Spherical directions and circular posterior draws}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 5)
```

# Spherical helpers

The spherical helpers convert between azimuth-colatitude coordinates and
Cartesian unit vectors. They are intentionally lightweight and are designed for
preprocessing or summary steps.

```{r}
library(ggplot2)
library(ggcircular)

coords <- spherical_to_cartesian(
  theta = c(0, pi / 2, pi),
  phi = c(pi / 2, pi / 2, pi / 3)
)

coords
cartesian_to_spherical(coords$x, coords$y, coords$z)
spherical_summary(c(0, pi / 2, pi), c(pi / 2, pi / 2, pi / 3))
```

# Circular posterior draws

Posterior draws can be converted to a long circular format when the optional
`posterior` package is installed.

```{r}
if (requireNamespace("posterior", quietly = TRUE)) {
  set.seed(1)
  draws <- posterior::draws_df(
    theta = rnorm(400, mean = pi / 3, sd = 0.25),
    phi = rnorm(400, mean = pi, sd = 0.35)
  )

  circular_draws <- as_circular_draws(draws, variables = c("theta", "phi"))
  summarise_circular_draws(circular_draws)
}
```

```{r}
if (requireNamespace("posterior", quietly = TRUE)) {
  autoplot_circular_draws(circular_draws)
}
```

# State-based angle plots

The same plotting grammar applies when angles are grouped by observed or latent
states.

```{r}
plot_state_angles(animal_steps, angle = turn_angle, state = state, type = "rose")
```
