---
title: "Comparative validation notes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Comparative validation notes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 5)
```

# Scope

This article records lightweight comparative checks for `ggcircular`. It is a
pkgdown article rather than a CRAN vignette because it may grow into a longer
validation document for a software paper.

The focus is practical agreement on core quantities:

1. circular mean direction;
2. mean resultant length;
3. axial summaries;
4. von Mises simulation and mixture diagnostics;
5. timing and memory proxies for common plotting workflows.

```{r}
library(ggplot2)
library(dplyr)
library(ggcircular)
```

# Optional package availability

```{r}
comparison_packages <- c("circular", "CircStats", "NPCirc", "Directional")

tibble(
  package = comparison_packages,
  available = vapply(comparison_packages, requireNamespace, logical(1), quietly = TRUE)
)
```

# Boundary behaviour

```{r}
boundary <- tibble(theta = c(0.02, 0.04, 2 * pi - 0.03, 2 * pi - 0.01))

boundary |>
  summarise(
    arithmetic_mean = mean(theta),
    circular_mean = mean_direction(theta),
    Rbar = mean_resultant_length(theta)
  )
```

```{r}
ggplot(boundary, aes(x = theta)) +
  geom_rose(bins = 16, alpha = 0.7) +
  geom_mean_direction(colour = "#123C4A", linewidth = 1) +
  scale_x_circular_degrees() +
  coord_circular() +
  theme_circular()
```

# Comparison with circular when available

```{r}
if (requireNamespace("circular", quietly = TRUE)) {
  x <- c(rnorm(120, 0.5, 0.25), rnorm(120, 5.8, 0.25))
  x <- normalize_angle(x)

  circular_x <- circular::circular(x)

  tibble(
    statistic = c("mean", "Rbar"),
    ggcircular = c(mean_direction(x), mean_resultant_length(x)),
    circular = c(
      as.numeric(circular::mean.circular(circular_x)),
      as.numeric(circular::rho.circular(circular_x))
    ),
    absolute_difference = abs(ggcircular - circular)
  )
}
```

# Uniform data

```{r}
uniform_angles <- seq(0, 2 * pi, length.out = 181)[-181]

tibble(
  mean = mean_direction(uniform_angles),
  Rbar = mean_resultant_length(uniform_angles),
  rayleigh_p_value = rayleigh_test(uniform_angles)$p.value
)
```

# Axial data

```{r}
axial_example <- tibble(
  orientation = normalize_angle(c(
    rnorm(60, 0.1, 0.08),
    rnorm(60, pi + 0.1, 0.08)
  ))
)

axial_example |>
  summarise(
    directional_Rbar = mean_resultant_length(orientation),
    axial_Rbar = mean_resultant_length(orientation, axial = TRUE),
    axial_mean = mean_direction(orientation, axial = TRUE)
  )
```

```{r}
ggplot(axial_example, aes(x = orientation)) +
  geom_rose(bins = 18, axial = TRUE, alpha = 0.75) +
  geom_mean_direction(axial = TRUE, colour = "#123C4A", linewidth = 1) +
  scale_x_circular_degrees(limits = c(0, pi)) +
  coord_circular() +
  theme_circular()
```

# Bimodal mixture

```{r}
set.seed(2026)
bimodal <- normalize_angle(c(rnorm(100, 0.8, 0.2), rnorm(100, 4.1, 0.25)))
fit <- fit_vonmises_mixture(bimodal, k = 2, nstart = 5, seed = 2026)

tidy_circular(fit)
glance_circular(fit)
```

```{r}
ggplot(tibble(theta = bimodal), aes(x = theta)) +
  geom_rose(aes(y = after_stat(density)), bins = 24, alpha = 0.45) +
  stat_vonmises_mixture(fit = fit, linewidth = 1) +
  scale_x_circular_degrees() +
  coord_circular() +
  theme_circular()
```

# Timing proxy

```{r}
set.seed(1)
timing_data <- normalize_angle(rnorm(1000, 1, 0.7))

system.time({
  density_tbl <- ggplot_build(
    ggplot(tibble(theta = timing_data), aes(x = theta)) +
      stat_circular_density(n = 256)
  )$data[[1]]
})

density_tbl |>
  summarise(n_grid = n(), density_min = min(density), density_max = max(density))
```

# Applied examples to keep validating

The main validation examples for an article should remain reproducible and
small:

1. wind direction with compass bearings;
2. animal movement turn angles by state;
3. axial orientation data;
4. circular residual diagnostics from angular regression;
5. posterior angular draws when `posterior` is installed.

These examples are already represented in the package datasets and focused
articles. The next validation pass should add formal numerical comparisons for
packages that are installed in the full-suggests CI profile.
